Summary of PM2.5 Nonattainment  Analysis

### 2020-11-08 Analysis Files/Paper_Analysis_327.Rmd
added random forest prop score moethod

```{r}
rm(list=ls())
set.seed(51)
# library(HEIfunctions)
library(maps)
library(PSAgraphics)
library(data.table)

@@

### Basic Data Description (Section 2.2)
```{r,  echo=FALSE}
print(paste("Total number of locations =", dim(dat)[[1]]))
print(paste("Total number of fee-for-service Medicare beneficiaries =", sum(dat$Total_den_FFS.2012)))
print(paste("Total number of managed care Medicare beneficiaries =", sum(dat$Tot_den_for_death_MA_FFS.2012) - sum(dat$Total_den_FFS.2012)))
```
### Nonattainment Description (Section 2.2)
```{r,  echo=FALSE}
print(paste("Number of nonattainment areas =", sum(dat$a==1)))
print(paste("Number of nattainment areas =", sum(dat$a==0)))

whichyears = 1999:2012  ## only 3 monitors in 1997, only 4 in 1998
whichvar = "Arithmetic.Mean" ##"X95th.Percentile" ##Arithmetic.Mean"
attaindat=read.csv('Data Construction/Other Data Files/EPA Nonattainment Table.csv')
attaindat$FIPS=paste(formatC(attaindat$fips_state, width=2, flag="0"), formatC(attaindat$fips_cnty, width=3, flag="0"), sep="")
attaindat = subset(attaindat, pollutant=='PM-2.5 1997')
attaindat$a_2012 = ifelse(attaindat$pw_2012 %in% c("P", "W"), 1, 0)
checka_2012 = attaindat[, c("FIPS", "a_2012")]
check_attain_fu = merge(dat, checka_2012, by = "FIPS", all.x = TRUE)
check_attain_fu$a_2012[is.na(check_attain_fu$a_2012)] = 0
with(check_attain_fu, table(a, a_2012))
print(paste("Percentage of 2005 nonattainment areas that remained in nonattainment in 2012 =", with(check_attain_fu, round(sum(a_2012==1)/(sum(a==1)),2))))

@@

# Propensity Score Analysis
```{r, echo=FALSE}
###############################################################
#####     Estimate the Propensity Score      ##################
###############################################################
dat$CompletelyRural = as.numeric(dat$CompletelyRural)
# -- Full model using only summary health outcomes
fullmodel = glm(a~.  + avgrelhum*avgtemp + pmbase*avgtemp, family=binomial(link="logit"), data = dat[, names(dat) %in% c(covariates, "a") & !(names(dat) %in% c("Monitor", "TotPop", "COPDrate", "CVstrokerate", "HFrate", "HRDrate", "IHDrate", "PVDrate", "RTIrate")), with = FALSE]) #"mortrate", "resprate", "CVDrate"

# - Take out Census covariates redundant with Medicare
allvars = rownames(summary(fullmodel)$coef)
inclvars = allvars[!(allvars) %in% c("PctBlack", "PctHisp", "PctFemale")]
inclvars = inclvars[!inclvars %in% c("(Intercept)")]
inclformula = as.formula(paste("a ~", paste(inclvars, collapse = "+")))
psmod = glm(inclformula, family=binomial(link="logit"), data = dat) 
print("The propensity score model summary:")
summary(psmod)

dat$ps = round(psmod$fitted.values, 3)
mina1 = min(dat$ps[dat$a==1])
maxa0 = max(dat$ps[dat$a==0])
dat$outofrange = 0

>>>>

```{r}
rm(list=ls())
set.seed(51)
# library(HEIfunctions)
library(maps)
library(PSAgraphics)
library(data.table)
library(randomForest)

@@

### Basic Data Description (Section 2.2)
```{r,  echo=FALSE}
print(paste("Total number of locations =", dim(dat)[[1]]))
print(paste("Total number of fee-for-service Medicare beneficiaries =", sum(dat$Total_den_FFS.2012)))
print(paste("Total number of managed care Medicare beneficiaries =", sum(dat$Tot_den_for_death_MA_FFS.2012) - sum(dat$Total_den_FFS.2012)))
```
### Nonattainment Description (Section 2.2)
```{echo=FALSE}
print(paste("Number of nonattainment areas =", sum(dat$a==1)))
print(paste("Number of nattainment areas =", sum(dat$a==0)))

whichyears = 1999:2012  ## only 3 monitors in 1997, only 4 in 1998
whichvar = "Arithmetic.Mean" ##"X95th.Percentile" ##Arithmetic.Mean"
attaindat=read.csv('Data Construction/Other Data Files/EPA Nonattainment Table.csv')
attaindat$FIPS=paste(formatC(attaindat$fips_state, width=2, flag="0"), formatC(attaindat$fips_cnty, width=3, flag="0"), sep="")
attaindat = subset(attaindat, pollutant=='PM-2.5 1997')
attaindat$a_2012 = ifelse(attaindat$pw_2012 %in% c("P", "W"), 1, 0)
checka_2012 = attaindat[, c("FIPS", "a_2012")]
check_attain_fu = merge(dat, checka_2012, by = "FIPS", all.x = TRUE)
check_attain_fu$a_2012[is.na(check_attain_fu$a_2012)] = 0
with(check_attain_fu, table(a, a_2012))
print(paste("Percentage of 2005 nonattainment areas that remained in nonattainment in 2012 =", with(check_attain_fu, round(sum(a_2012==1)/(sum(a==1)),2))))
```

@@

# Propensity Score Analysis - Updated to include Random Forest!
```{r, echo=FALSE}
###############################################################
#####     Estimate the Propensity Score      ##################
###############################################################
dat$CompletelyRural = as.numeric(dat$CompletelyRural)
# -- Full model using only summary health outcomes
fullmodel = glm(a~.  + avgrelhum*avgtemp + pmbase*avgtemp, family=binomial(link="logit"), data = dat[, names(dat) %in% c(covariates, "a") & !(names(dat) %in% c("Monitor", "TotPop", "COPDrate", "CVstrokerate", "HFrate", "HRDrate", "IHDrate", "PVDrate", "RTIrate")), with = FALSE]) #"mortrate", "resprate", "CVDrate"

# - Take out Census covariates redundant with Medicare
allvars = rownames(summary(fullmodel)$coef)
inclvars = allvars[!(allvars) %in% c("PctBlack", "PctHisp", "PctFemale")]
inclvars = inclvars[!inclvars %in% c("(Intercept)")]
inclformula = as.formula(paste("a ~", paste(inclvars, collapse = "+")))

get_propensity_scores <- function(method="logit") {
  if (method == "logit") {

    psmod = glm(inclformula, family=binomial(link="logit"), data = dat) 
    print("The propensity score model summary:")
    summary(psmod)
    return(round(psmod$fitted.values, 3))

  } else if (method == "RF") {

    inclvars_rf = inclvars[!inclvars %in% c("avgtemp:pmbase", "avgtemp:avgrelhum")]
    dat_rf <- dat[, names(dat) %in% c(inclvars_rf, "a"), with = FALSE]
    dat_rf$avgtemp_avgrelhum <- dat_rf$avgrelhum * dat_rf$avgtemp
    dat_rf$avgtemp_pmbase <- dat_rf$pmbase * dat_rf$avgtemp

    fit_rf <- randomForest(x = dat_rf[,!"a"], y = as.factor(dat_rf$a))
    return(fit_rf$votes[,2])

  }
}

# dat$ps = get_propensity_scores()
dat$ps = get_propensity_scores(method="RF")

mina1 = min(dat$ps[dat$a==1])
maxa0 = max(dat$ps[dat$a==0])
dat$outofrange = 0
